---
title: "Minimal Example"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Minimal Example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


# Intro

This vignette is a short explanation of the structure of the `CI.RCT.Sim`
package. And gives an explanation on how to contribute code for the
data-generation of scenarios as well as for analysis functions.

The example uses a simple example of a test and an estimator for the difference
in means of two normal distributions.

For to run the simulations the package `SimDesign` is used. Some functions for
sumarisation of simulation results are re-used from the `SimNPH` package written
for a previous simulation study.

# Data Generation

Let's look at the code for the data generation. The simulation study uses the
`SimDesign` pacakge. To run simulations with SimDesign, one has to define three
functions for *Data Generation*, *Analysis* and *Summarisation*. Each of those
functions has to use fixed parameters. The function signatures can be displayed
with `SimDesign::SimFunctions()`. This is the first function
`generate_minimal_example` displayed in the code below.

In `CI.RCT.Sim` there's two additional functions for each data-generating
function. `true_summary_statistics_minimal_example`, calculates the true (or
assymptotic) values of summary statistics from the parameters given in a
"Design" dataset. This is specific to the data generating model and the
estimand.

The third function is a function that prints and outputs an example of a Design
dataset that can be used with the data generating functions. This function gives
a dataset that is directly useable with the data generating function for
testing, documentation of the parameter and for users to have boilerplate code
to be able to cerate custom parameter ranges.

`R/generate_minimal_example.R`
```{r, file="../R/generate_minimal_example.R", echo=TRUE, eval=FALSE}
```

The true summary statistics is really simple here, it is just the difference of
group means.

The utility function `params_scenario_grid` expands a dataset in the following
way: the first row is the reference scenario with the first value of each
defined parameter. The subsequent rows contain scenarios in which each parameter
is varied across all given values and all other parameters are kept at the
reference scenario value.

# Analysis

The analysis functions used in `SimDesign` take the arguments `condition` and
`dat`, additional arguments can be passed in the optional argument
`fixed_objects`. When running the simulations the current line of the Design
dataset, containing the parameters for the current simulation scenario is passed
as the `condition` argument. The simulated dataset is passed as the `dat`
argument. Each analysis function is called once for each simulation replication
for each scenario.

To be able to use hyper-parameters for the analysis methods, CI.RCT.Sim
implements the analysis functions such that they are functions with the
hyper-parameters as arguments that return a function that can be used in
`SimDesign::RunSimulation`. This way one analysis method can easily be used
multiple times with different hyper-parameters in one simulation.

For the minimal example two analysis functions are defined,
`analyse_minimal_example_lm` uses a linear model and `analyse_minimal_example_t`
uses a t-test. The return value for each method should be a named list to enable
summarisation later.

`R/analyse_minimal_example.R`
```{r, file="../R/analyse_minimal_example.R", echo=TRUE, eval=FALSE}
```

Some names of the names of the analysis results are directly used by analysis
functions, others can be choosen freely. "reserved" names are:

* `p` the p-value of tests
* `N_pat` number of recruited patients, relevant for sequential or adaptive designs
* `N_evt` number of observed events, relevant for time-to event outcomes under censoring

# Running Simulations

To actually run the simulations, first a dataset of simulation parameters is
created and true values of summary statistics are calculated if necessary. A
list of analysis functions is created, here hyper-parameters are set, in this
example the coverage of the confidence intervals.

A list of summary functions is defined, those analysis functions are the applied
to the list of outputs of all replications of each scenario. Which summary
functions are applied to the output of which analysis functions is matched by
name by the function `SimNPH::create_summary_function`, names can be repeated in
the list of summary functions. The functions `SimNPH::summarise_estimator` and
`SimNPH::summarise_test` provide common summaries for estimators and tests, see
the documentation in the SimNPH package for details.

Finally `SimDesign::runSimulation` is used to conduct the actual simulations.
This function takes care of parallelisation, initialisation of random seeds etc.

`vignettes/minimal_example.R`
```{r, file="../vignettes/minimal_example.R", echo=TRUE}
```
